### 2022-01-26 article_issac.tex
langage régulier

\documentclass[sigconf]{acmart}

\copyrightyear{2021}
\acmYear{2021}
\setcopyright{rightsretained}
\acmConference[ISSAC '22]{International Symposium on Symbolic and Algebraic Computation}{July 4--7, 2022}{Lille, France}
\acmBooktitle{International Symposium on Symbolic and Algebraic Computation (ISSAC '22), July 4--7, 2022, Lille, France}
\acmPrice{15.00}
\acmDOI{XX.XXX/XXXXXX.XXXXXX}
\acmISBN{XXXXXXXXXXXXXXXXXXX}

\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%\usepackage{lmodern}
%\usepackage[a4paper]{geometry}

\usepackage{amsmath,  mathrsfs}
\usepackage[all]{xy}
\usepackage{stmaryrd}

\usepackage{tikz}
\usepackage{hyperref}

\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%\usepackage{graphicx}

\usepackage[english]{babel}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{lem}{Lemma}[section]
\newtheorem{prop}[lem]{Proposition}
\newtheorem{thm}[lem]{Theorem}
\newtheorem{cor}[lem]{Corollaire}
\newtheorem{axi}{Axiome}

\theoremstyle{definition}
\newtheorem{defn}[lem]{Definition}

\theoremstyle{remark}
\newtheorem{step}{\'Etape}
\newtheorem*{rmk}{Remark}
\newtheorem*{expl}{Example}

\newcommand{\N}{\mathbb{N}}
\newcommand{\NN}{\mathbb N}
\newcommand{\MM}{\mathbb M}

\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Zp}{\mathbb{Z}_p}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Qp}{\Q_p}
\newcommand\cO{\mathcal{O}}
\newcommand{\softO}{O\tilde{~}}
\newcommand{\ddx}{\frac{\mathrm{d}\phantom{x}}{\mathrm{d}x}}
\newcommand{\OKun}{O_K([0,1])}
\newcommand{\val}{\mathrm{val}}

\newcommand\cY{\mathcal{Y}}
\newcommand{\ud}{\mathrm{d}}

\newcommand{\LT}{LT}
\newcommand{\LC}{LC}
\newcommand{\LM}{LM}

\newcommand\Zpt{\Zp\llbracket t \rrbracket}
\newcommand\Qpt{\Qp\llbracket t \rrbracket}
\newcommand\wrt{with respect to\xspace}

\newcommand{\X}{\mathbf{X}}
\newcommand{\Y}{\mathbf{Y}}
\renewcommand{\i}{\mathbf{i}}
\renewcommand{\j}{\mathbf{j}}
\renewcommand{\r}{\mathbf{r}}

\newcommand\fixmecc[1]{{\color{red}{\underline{\bf C:}} #1}}

% Macros report
\newcommand{\M}{\mathcal{M}}

\newcommand{\ifnonempty}[3]{%
  % IF #1 is empty THEN #3 ELSE #2
  \def\tempa{}%
  \def\tempb{#1}%
  \ifx\tempa\tempb % Empty case
  #3 
  \else            % Non-empty case
  #2
  \fi}

\newcommand{\KX}{K \left\langle X \right\rangle}

\def\todo#1{\ \!\!{\color{red} #1}}

\def\eqdef{\stackrel{\text{def}}{=}}

\begin{document}

\fancyhead{}

\title{On The Difficulty of Computing Non-Commutative Signature Gröbner Bases}

\author{Cyrille Chenavier}
\affiliation{Universit\'e de Limoges;
  \institution{CNRS, XLIM UMR 7252}
  \city{Limoges}
  \country{France}  
  \postcode{87060}  
}
\email{cyrille.chenavier@unilim.fr}

\author{
  Arthur Léonard}
\affiliation{\'Ecole Normale Supérieure, 
  \institution{DI, DMA}
  \city{Paris}
  \country{France}  
}
\email{arthur.leonard@ens.psl.eu }

\author{
  Tristan Vaccon}
\affiliation{Universit\'e de Limoges;
  \institution{CNRS, XLIM UMR 7252}
  \city{Limoges}
  \country{France}  
  \postcode{87060}  
}
\email{tristan.vaccon@unilim.fr}

\thanks{This work was supported by...}

\begin{abstract}
NC GB, signature...\end{abstract}

\begin{CCSXML}
  <ccs2012>
  <concept>
  <concept_id>10010147.10010148.10010149.10010150</concept_id>
  <concept_desc>Computing methodologies~Algebraic algorithms</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Algebraic algorithms}

% \vspace{-1mm}
% \ccsdesc[500]{Computing methodologies~Algebraic algorithms}
% \printccsdesc

\vspace{-1.5mm}
\terms{Algorithms, Theory}

\keywords{Algorithms, Gröbner bases, non-commutative algebra}

\maketitle

%\clubpenalty=10000
\widowpenalty = 10000
\addtolength{\textfloatsep}{-0.45cm} % Distance between float (e.g. [t]) and text

\section{Introduction}
Commutative part.
Modern algorithms:
\cite{F99,F5,FGLM}.
Survey \cite{EF17}
New take  \cite{G2V, GVW}
Applications \cite{LWXZ, CVV20, VVY21}

NC part.
Intro \cite{Mora94}
Freaks \cite{GMU00}.
F5 for right-modules over path algebra quotients \cite{K14}.
GVW for skew polynomial rings \cite{ZZ15}
Final complete
nc version \cite{HV21}.

\section{Notations}

In this article,
we work with
the notations defined
in \cite{HV21}.

%Aller jusqu'à la Def 18
%de l'article HV

Let $X= \{ x_1,\dots,x_n \}$
be a set of indeterminates.
We denote by $\left\langle X \right\rangle$ the
free monoid of words 
over $X$ and, for a field
$K$, we denote by
$\KX$ the
free algebra generated by $X$
over $K.$
We consider the elements
in $\KX$
as noncommutative polynomials
with coefficients in $K$
and indeterminates in $X$
(indeterminates commute
with coefficientes but not
with each other).

If $F \subset \KX,$
we denote by $(F)$
the two-sided ideal 
generated by $F$:
\[(F) = \left\lbrace \sum_{i=1}^d a_i f_i b_i \: \mid \: f_i \in F, \: a_i,b_i \in \KX, \: d \in \NN  \right\rbrace. \]

Let $\leq$ be a monomial 
ordering on $\left\langle X \right\rangle$.
For $f \in \KX \setminus \{ 0 \}$, we denote by $\LC (f), \LT (f), \LM (f)$
the leading coefficients,
leading terms and 
leading monomials of $f$
(with $\LT(f) = \LC(f) \LM(f)$).

We recall that $G \subset (F)$
is a Gröbner basis
of $(F)$ if
$\LM (G)$ generates $LM\left( (F) \right)$ as
monoids inside $\left\langle X \right\rangle.$

For $r \in \NN$ we denote
by $\mathscr{F}_r = \left( \KX \otimes \KX \right)^r$
the free $\KX$-bimodule
of rank $r$, with canonical
basis $\varepsilon_1,\dots,\varepsilon_r$
where $\varepsilon_i =(0,\dots,0,1 \otimes 1,0,\dots,0)$ ($1 \otimes 1$ in the $i$-th position) for $i \in \llbracket 1,r \rrbracket.$
Let $\MM \left( \mathscr{F}_r \right)= \left\lbrace a \varepsilon_i b \mid a,b \in \left\langle X \right\rangle, \: i \in \llbracket 1,r\rrbracket \right\rbrace$
be the set of module monomials in $\mathscr{F}_r.$ It is a $K$-vector space basis
of $\mathscr{F}_r.$
Let $\leq_\MM$ be a module
ordering on $\MM \left( \mathscr{F}_r \right).$
We recall that it means that
$\leq_\MM$ is a well-ordering
compatible with 
scalar multiplication, that
is 
$\mu \leq_\MM \mu'$ implies
$a \mu b \leq_\MM a \mu' b$
for all $\mu,\mu' \in \MM \left( \mathscr{F}_r \right)$
and $a, b \in \left\langle X \right\rangle.$
We assume that 
$\leq$ and $\leq_\MM$
are compatible
(\textit{i.e.} for any
$a,b \in \left\langle X \right\rangle$, $i \in \llbracket 1,r \rrbracket$,
$a \leq b \Leftrightarrow a \varepsilon_i \leq_\MM b \varepsilon_i \Leftrightarrow \varepsilon_i a \leq_\MM \varepsilon_i b$).
We also assume that 
$\leq_\MM$ is fair (for any
$\mu \in \MM (\mathscr{F}_r)$,
the set $\left\lbrace \mu' \in \MM (\mathscr{F}_r) \mid \mu' \leq_\MM \mu \right\rbrace$ is finite).

If $\alpha \in \mathscr{F}_r$, it can be uniquely
written in the basis 
$\MM \left( \mathscr{F}_r \right) $
and the biggest monomial
involved is its signature
$\mathfrak{s}(\alpha) \in
\mathscr{F}_r$.
\fixmecc{plutôt $\mathfrak{s}(\alpha) \in \MM (\mathscr{F}_r)$? Aussi définir
$\mathfrak{s}(E)$ pour un sous ensemble $E$?}

If $M \subset \mathscr{F}_r$
is a $\KX$-submodule,
then $G \subset M$ is called
a Gröbner basis of $M$ if
\[\mathfrak{s}(M)= \left\lbrace a \mathfrak{s}(\gamma) b \mid a,b \in \left\langle X \right\rangle, \gamma \in G \right\rbrace. \]

Let $f_1,\dots,f_r \in \KX$
be some polynomials
generating the ideal $I=(f_1,\dots,f_r).$
They define a 
$\KX$-module homomorphism:

\[
\begin{array}{ccccc}
&& \mathscr{F}_r & \rightarrow & \KX \\
\overline{\cdot}& : &\alpha = \sum_i c_i a_i \varepsilon_{j_i} b_i&\mapsto& \overline{\alpha}=\sum_i c_i a_i f_{j_i} b_i. \\
\end{array}\]

Elements $(f,\alpha) \in \KX \times \mathscr{F}_r$
such that $f=\overline{\alpha}$
are called signature polynomials.
We denote by $f^{[\alpha]}$
such a pair. 
For such a pair, necessarily,
$f \in I.$
We denote by $I^{[\Sigma ]}=\left\lbrace f^{[\alpha]} \mid f=\overline{\alpha} \right\rbrace \subset I \times \mathscr{F}_r$
the set of all signature
polynomials.

For $f^{[\alpha]},g^{[\beta]},
h^{[\gamma]} \in I^{\Sigma},$
with $\alpha, g \neq 0,$
we say that 
$f^{[\alpha]}$ $\mathfrak{s}$-reduces to
$g^{[\beta]}$ by
$h^{[\gamma]}$
if there exist $a,b \in 
\left\langle X \right\rangle$
such that:
\begin{itemize}
\item $a \LM(h)b \in supp(f),$
\item $\mathfrak{s}(a \gamma b) \leq \mathfrak{s}(\alpha),$ and
\item $g^{[\beta]}=f^{[\alpha]}-\frac{coeff(f,a \LM(h)b}{\LC(g)} a h^{[\gamma]}b.$
\end{itemize}
Reduction by a set $G^{[\Sigma]} \in I^{[\Sigma]}$
is then defined as the 
reflexive, transitive closure
of the reduction by elements
of $G^{[\Sigma]} $.

Finally, $G^{[\Sigma]} $
is called a \textit{signature
Gröbner basis} of $I^{[\Sigma]}$ if all
$f^{[\alpha]} \in I^{[\Sigma]}$ $\mathfrak{s}$-reduce to zero by $G^{[\Sigma]}.$

\section{Rationality}

Looking at principal syzygies and the algorithm for computing new ones, we wonder if an automaton can recognize leading monomials of syzygies in the case where the Gröbner basis is finite. Let us show that this is not the case using the non regular language\linebreak $\{a^nb^n \;|\; n \in \N\}$. Let us consider the alphabet 
$$\{a, b, c, d, e, s, t, x, y, z, \infty\}$$
and polynomials
\begin{align*}
g_1 & := sa - tcx \\
g_2 & := xa - ax \\
g_3 & := xd - dx \\
g_4 & := ya - ay \\
g_5 & := yd - dy \\
g_6 & := xbb - dyb \\
g_7 & := cya - ccx \\
g_8 & := xbe - dze \\
g_9 & := dz - zb \\
g_{10} & := cz - za \\
g_{11} & := tz - s
\end{align*}
together with the order $m_1 < m_2$ if:
\begin{itemize}
\item there are less $\infty$ symbols in $m_1$ than in $m_2$,
\item $m_1$ and $m_2$ have the same number of $\infty$ symbols and there are less $s$ symbols in $m_1$ than in $m_2$,
\item $m_1$ and $m_2$ have the same numbers of $\infty$ and $s$ symbols, but
$m_1 <_{deglex} m_2$ where\\ $a < b < c < d < x < y < z < e < t < s < \infty$.
\end{itemize}

\fixmecc{Given $l,r,l',r'\in\KX$, have (let?) $l\varepsilon_i r\leq l'\varepsilon_j r'$ if
  \begin{itemize}
  \item $l\LM(g_i)r < l'\LM(g_j)r'$
  \item $l\LM(g_i)r = l'\LM(g_j)r'$ et $i < j$
  \item $l\LM(g_i)r = l'\LM(g_j)r'$ et $i = j$ et $\deg{l} < \deg{l'}$.
  \end{itemize}
  The idea is that $s, t, e$ symbols take place at the ends of a word in order to handle the following operations between the first $a$ and the last $b$.
  The first $a$ will transform and generate an $x$, other $a$'s will absorb a $y$, and generate an $x$ will transforming into a $c$.
The last $b$ will absorb an $x$, generate a $z$ while transforming into a $d$, other $b$'s will absorb an $x$, generate a $y$ while transforming into a $d$.
The $x$'s and $y$'s are messages which commute with $a$'s and $d$'s. A $x$ "dit" transmits the information to the border between $b$'s and $d$'s than an $a$ transformed into a $c$, while a $y$ transmits the information to the border between $a$'s and $c$'s than a $b$ transformed into a $d$. The objective being that in order to compare the numbers of $a$'s and $b$'s, there are the same number of transformations of each type.
The $z$ is the message which enables us to transform $d$'s into $b$'s and $c$'s into $a$'s at the end of the procedure.
We will explain latter the role played by $\infty$ symbol.
}

{\color{blue}{\bf\underline{Version française:}}  Nous avons $(l, i, r) \leq (l', j, r')$ si l'une des conditions ci-desous est vérifiée~:
\begin{itemize}
	\item $l\LM(g_i)r < l'\LM(g_j)r'$
	\item $l\LM(g_i)r = l'\LM(g_j)r'$ et $i < j$
	\item $l\LM(g_i)r = l'\LM(g_j)r'$ et $i = j$ et $\deg{l} < \deg{l'}$.
\end{itemize}

L'idée est que les deux symboles $s, t, e$ sont présents aux extrémités du mot pour pouvoir faire des opérations sur le premier $a$ ou sur le dernier $b$.
Le premier $a$ va se transformer en $c$ en générant un $x$, les autres $a$ vont absorber un $y$, et générer un $x$ en se tranformant en $c$.
Le dernier $b$ va absorber un $x$, générer un $z$ et se transformer en $d$, les autres $b$ vont absorber un $x$, et générer un $y$ en se transformant en $d$.
Les $x$ et les $y$ sont des "messages" qui commutent avec les $a$ et les $d$. Un $x$ "dit" à la frontière entre les $b$ et les $d$ qu'un $a$ s'est transformé en $c$, alors qu'un $y$ "dit" à la frontière entre les $a$ et les $c$ qu'un $b$ s'est transformé en $d$, l'objectif étant que pour comparer le nombre de $a$ et de $b$, il y ait autant de transformations de chaque type.
Le $z$ est le message qui permet de retransformer les $d$ en $b$ et les $c$ en $a$ à la fin du processus. On reparlera du rôle de $\infty$ plus tard.
}

For a better understanding, let us consider the case $n=2$. The monomial $\varepsilon_{11}a^2b^2e$ is a leading monomial od a syzygy because of the following sequence of equalities.
\begin{align*}
saabbe & =_{1} tcxabbe =_{2} tcaxbbe =_{6} tcadybe \\
& =_{5} tcaydbe =_{4} tcyadbe =_{7} tccxdbe \\
& =_{3} tccdxbe =_{8} tccddze =_{9} tccdzbe \\
& =_{9} tcczbbe =_{10} tczabbe =_{10} tzaabbe \\
& =_{11} saabbe
\end{align*}

\begin{proposition}
  Leading monomials of syzygies of $F=\{g_1,\dots,g_{11}\}$ do not form a regular language
\end{proposition}
\begin{proof}
  For every $n \geq 1$, $g_{11}a^nb^ne$ is the leading monomial of the syzygy:

  \begin{align*}
    g_1a^{n-1}b^ne & + \sum_{k = 2}^{n} tc^{k - 2}g_7a^{n - k}d^{k - 1}b^{n - k + 1}e \\
    & + \sum_{k = 1}^{n - 1} tc^ka^{n - k}d^{k - 1}g_6b^{n - k - 1}e \\
    & + c^nd^{n - 1}g_8 \\
    & + g_{11}a^nb^ne \\
    & + \sum_{k = 1}^n tc^nd^{n - k}g_9b^{k - 1}e \\
    & + \sum_{k = 1}^n tc^{n - k}g_{10}a^{k - 1}b^ne \\
    & + \sum_{k = 1}^{n} \sum_{l = 0}^{n - k - 1} tc^{k}a^{l}g_2a^{n - l - k - 1}d^{k - 1}b^{n - k + 1} e \\
    & + \sum_{k = 1}^{n} \sum_{l = 0}^{k - 2} tc^{k}a^{n - k}d^lg_3d^{k - 2 - l}b^{n - k - 1}e \\
    & + \sum_{k = 1}^{n - 1} \sum_{l = 0}^{k - 1} tc^{k}a^{n - k}d^lg_5d^{k - 1 - l}b^{n - k - 2}e \\
    & + \sum_{k = 1}^{n - 1} \sum_{l = 0}^{n - k - 1} tc^{k}a^{l}g_4a^{n - l - k - 1}d^{k}b^{n - k} e. \\
  \end{align*}

  Now, let us show that leading monomials of syzygies of the form $\varepsilon_{11}a^nb^me$ are such that $n = m$.
  Let $\sigma$ be such a syzygy and let $G$ be the graph whose vertices are elements of $\left\langle X \right\rangle$, and there is an edge between vertices of the form $lXr$ and $lYr$ as soon as $lg_ir$ occurs with a nonzero coefficient in $\sigma$ and $g_i=X - Y$. Let $W$ be the set of vertices of the following form:
  \begin{align*}
    & sa^nb^ne, \\
    & tc^pa^qxa^rd^nb^me, \\
    & tc^pa^qd^rxd^nb^me, \\
    & tc^pa^qya^rd^nb^me, \\
    & tc^pa^qd^ryd^nb^me, \\
    & tc^pza^qb^re, \\
    & tc^pd^qzb^re
  \end{align*}
  Each vertex in $W$ is related to at most two vertices, which also belong to $W$.
  Hence, the connected component of $sa^nb^me$ in $G$ is either a chain or a cycle (\fixmecc{je n'ai pas compris l'argument}).
  But since $\sigma$ is a syzygy, this connected component has to be a cycle, which implies $n = m$.

  Now, we can conclude that the set $X$ of leading monomials of syzygies of $F$ does not form a regular language.
  Indeed, by way of contradiction, assume that $X$ is a regular language, so that its
  intersection with the regular language $\{g_{11}a^nb^me \;|\; n, m \geq 1\}$ also forms regular language.
  We reach a contradiction since we proved that this intersection is $\{g_{11}a^nb^ne \;|\; n \geq 1 \}$, which does not form a regular language.
\end{proof}

Notice that $F=\{g_1,\dots,g_{11}\}$ has no reason to have a finite Gröbner basis. However, we construct a new set $F'$ by adding to $F$ the polynomial $\infty$ and all polynomials of the form $\infty-\alpha$, for letters $\alpha$ different from $\infty$. Then $F'$ admits the reduced Gröbner basis which contains each letter.
Now, let $X$ and $Y$ be the sets of leading monomials of syzygies of $F$ and $F'$, respectively, and let 
$$Z=\{l\varepsilon_ir \in \MM(\mathscr{F}_r) \;|\; l \LM(g_i) r \mbox{ does not contain } \infty\}.$$
Since $X=Y\cap Z$, $Z$ is a regular language and $Z$ is not a regular language, $Y$ is not a regular language. Hence, we have constructed a set $F'$ of polynomials that admit a finite Gröbner basis whose set of leading monomials of syzygies does not form a regular language. Moreover, we can make the order \fixmecc{lequel?} isomorphic to $\N$
by replacing $\infty$ and $s$  by $\infty^{10}$ and $s^5$, respectively car les degrés varient d'au plus une constante lors des réécritures contenues dans une syzygie \fixmecc{pas compris la justification :s}.

\section{Algebraicity}

\section{Finiteness}

\section{Implementation}

\bibliographystyle{plain}
\begin{thebibliography}{99}
  \renewcommand{\itemsep}{0em}
  % \renewcommand{\itemsep}{0.14em}

\bibitem{Bu65}
  Buchberger Bruno,
\newblock{{Ein Algorithmus zum Auffinden der Basiselemente des Restklassenringes nach einem nulldimensionalen Polynomideal (An Algorithm for Finding the Basis Elements in the Residue Class Ring Modulo a Zero Dimensional Polynomial Ideal)}},
\newblock English translation in J. of Symbolic Computation, Special Issue on Logic, Mathematics, and Computer Science: Interactions. Vol. 41, Number 3-4, Pages 475--511, 2006

  \bibitem{CVV20}
  Caruso, X., Vaccon T., Verron T.,
  \newblock {Signature-based algorithms for Gröbner bases over Tate algebras},
  \newblock {in Proceedings: ISSAC 2020, Kalamata, Greece.}

\bibitem{EF17}
Eder Christian and Faug{\`{e}}re Jean{-}Charles,
\newblock {A survey on signature-based algorithms for computing Gr{\"{o}}bner bases},
\newblock J. of Symbolic Computation, 2017

  \bibitem{F99}
  Faugère Jean-Charles,
  \newblock  {{A new efficient algorithm for computing Gröbner bases (F4)}},
  \newblock  {Journal of Pure and Applied Algebra}, {1999}

\bibitem{F5}
Faugère, Jean-Charles,
\newblock {A new efficient algorithm for computing Gröbner bases without reduction to zero (F5)},
\newblock{in Proceedings: {ISSAC'02}.}

  \bibitem{FGLM}
  Faugère, J.-C., Gianni, P., Lazard, D., Mora, T.,
  \newblock{Efficient computation of zero-dimensional Gr{\"o}bner bases by change of ordering},
  \newblock{J. of   Symbolic Computation 16~(4), 329--344, 1993}

  \bibitem{G2V}
Gao Shuhong, Guan Yinhua and Volny IV Frank,
\newblock{ A new incremental algorithm for computing Groebner bases},
\newblock{In Proceedings: {ISSAC'10}.}
%\newblock{ In Proceedings of the 2010 International Symposium on Symbolic and Algebraic Computation (ISSAC ’10). ACM, New York, NY, USA, 13–19.}

  \bibitem{GVW}
Gao Shuhong, Volny IV Frank, and Wang Mingsheng,
\newblock{ A new framework for computing Gröbner bases},
\newblock{ Mathematics of computation, 2016, vol. 85, no 297, p. 449-465.}

\bibitem{K14}
S.A. King,
\newblock {A non-commutative F5 algorithm with an application to the computation of Loewy layers},
\newblock {J. of Symbolic Computation} 65 (2014), pp. 111-129

  \bibitem{LWXZ}
  Lu Dong, Wang Dingkang, Xiao Fanghiu, Zhou Jie,
  \newblock  {{Extending the GVW Algorithm to Local Ring}},
  \newblock {Proceedings of 43th International Symposium on Symbolic and Algebraic Computation, {ISSAC}'18, New York, USA}

\bibitem{GMU00}
E.Green, T. Mora, V.Ufnarovski,
\newblock The Non-Commutative Gröbner Freaks.
\newblock Symbolic Rewriting Techniques. Vol. 15.

\bibitem{HV21}
C. Hofstadler, T. Verron,
\newblock {
Signature Gröbner bases, bases of syzygies and cofactor reconstruction in the free algebra}, 
\newblock  	arXiv:2107.14675, 2021

\bibitem{Mora94}
T. Mora,
\newblock {
An  introduction to  commutative 
and noncommutative Gröbner 
bases },
\newblock Theoretical Computer Science 134, 1994.

  \bibitem{Sage}
  \newblock {{S}ageMath, the {S}age {M}athematics {S}oftware {S}ystem ({V}ersion 9.2)}, The Sage Development Team, 2020, \url{http://www.sagemath.org}

\bibitem{VVY21}
T.Vaccon, T.Verron, K.Yokoyama,
\newblock
{On affine tropical F5 algorithms},
\newblock {
Journal of Symbolic Computation},
Volume 102,
2021.

\bibitem{ZZ15}
X. Zhao, Y. Zhang,
\newblock {A signature-based algorithm for computing Gröbner-Shirshov bases in skew solvable polynomial rings}
\newblock Open Mathematics, Volume: 13, Issue: 1, 2015.

\end{thebibliography}

\end{document}